Investigating soundscapes perception through acoustic scenes simulation.

### 2017-08-24 lafaySimulatedScenePerception.tex
toto

\documentclass[preprint,12pt]{elsarticle}
\usepackage{url}
\usepackage[utf8]{inputenc}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{caption}
\usepackage{sectsty}
\usepackage{amsmath}

@@

\newcommand{\gl}[1]{\textcolor{blue}{GL : #1}}
\newcommand{\nm}[1]{\textcolor{magenta}{#1}}
\newcommand{\ml}[1]{\textcolor{red}{#1}}
\newcommand{\jfp}[1]{\textcolor{green}{JFP : #1}}
\newcommand{\ie}{\emph{i.\,e.}}
\newcommand{\Ie}{\emph{I.\,e.}}
\newcommand{\cf}{cf.}
\newcommand{\Cf}{Cf.}
\newcommand{\vs}{\emph{vs.}}
\newcommand{\Vs}{\emph{Vs.}}
\newcommand{\al}{\emph{et~al.}}
\newcommand{\eg}{\emph{e.\,g.}}
\newcommand{\Eg}{\emph{E.\,g.}}
\newcommand{\etc}{\emph{etc.}}
\newcommand{\myfloatalign}{\centering}

@@

\begin{itemize}
\item \emph{checking the pleasantness of the simulated sound scene}: to check that the emotional difference produced by the i- and ni-scenes is also present in terms of perceived pleasantness, we look at the existence of differences between the average $\mathcal{A}_{scene}$ and the average $\mathcal{A}_{subject}$ for each type of scenes (i or ni). The significance is evaluated by a Student test on two independent populations for $\mathcal{A}_{scene}$ and by a Student test on two paired populations for $\mathcal{A}_{subject}$;
\jfp{ce qu'est Ascene dans ce paragraphe n'est pas clair :
A scene est défini comme la moyenne des scores d'une scène donnée pour tous les participants.

le test, lui, porte sur la moyenne des Ascene pour les deux catégories
même chose pour les sujets}
\item \emph{analysis of the structural features}: in order to evaluate whether the affective difference between  the i- and ni-scenes has a significative influence  on the structural features of the scenes \ie~ if significative differences exist between structural features and perceived pleasantness. The significance is measured with a Student test on two populations \jfp{je ne comprends pas bien ce test}
\item \emph{study of the influence of the structural features on the perceived pleasantness}: in order to evaluate a possible impact of the structural features on the perceived pleasantness, linear correlations between these features and $\mathcal{A}_{scene}$ are considered. The Pearson correlation coefficient is used for that purpose, on the basis of the standard methodology used in multidimensional studies;
\item \emph{analysis of the semantic features}: in order to determine if the affective difference between  the i- and ni-scenes has an influence on the scenes composition in terms of sound sources, or if specific classes are more frequently used to simulate a given environment (i or ni), a V-test is considered. Testing is done at every level of abstraction, and separately for texture and event classes. For each class $j$, and each environment type $k$ ($k=\lbrace i,ni\rbrace$), the value $V_{jk}$ is computed as follows:

@@

where $c$ is the number of available classes, $c_k$ the number of classes for a given type of environment $k$, $c_j$ the number of  $j$ classes, and $c_{jk}$ the number of $j$ classes used for an environment $k$. \jfp{il faudrait donner une référence pour ce test : bouquin Lebart, Morineau, Piron, STATISTIQUE EXPLORATOIRE MULTIDIMENSIONNELLE
peut être donner les bornes de variations de Vjk pour aider le lecteur à comprendre cet indicateur} \ml{y pas une ref en Anglais ?} The V-test tests the null hypothesis that the ratio $\frac{c_{jk}}{c}$ is not significatively different from the ratio $\frac{c_{jk}}{c_k}$. For a given environment $k$, and class $j$, if the hypothesis is rejected, the class $j$ is said to be typical with respect to environment $k$. The typical classes are called \textbf{sound markers};\jfp{on ne comprend pas bien l'hypothèse alternative du test.

VOIR BOUQUIN LEBART page 184
ou bien article de Morineau sur le sujet

Vjk est une estimation approchée de la valeur test
}

@@

\begin{table*}[t]
 \setlength{\tabcolsep}{0.2pt}
 \centering
  {\renewcommand{\arraystretch}{0.9}
\begin{tabular}{c c c c}
Abstraction        & \multicolumn{2}{c}{Event sound markers} \\
level & i-scenes & ni-scenes \\
\hline
0  &                          & construction work (3.78)  \\
\hline
  & church bell  (4.5)             & car horn  (3.9) \\
1 & bicycle bell  (4.3)      & siren (3.9)\\
  & animal (4.2)              &       \\
   \hline
  & bird        (4.8)       & car horn  (4.0)\\
2 & church bell  (4.4)             & siren (4.0)\\
  & bicycle bell     (4.2)             &       \\
   \hline
  & bird song (4.8)        & klaxon  (4.1)\\
3 & church bell   (4.3)            & siren (4.0)\\
  & bicycle bell      (4.2)   &       \\
  & foot steps  (3.6)      &  \\
  &                           & \\
  & \multicolumn{2}{c}{Texture sound markers}      \\
  & i-scenes & ni-scenes \\
\hline
0 &     courtyard / park (4.1) &  construction (3.9)  \\
\hline
1 &     park (3.65)          &  crossroad (3.6)  \\
  &                          &  construction vehicule (3.3)  \\
\hline
2 &     park (3.64)          &  crossroad (3.56)  \\
\hline
\end{tabular}
}
\vspace{0.5mm}
\caption{Event and texture classes identified as sound markers. In each cell, markers are ranked in the decreasing order of V-test value.\jfp{a partir de quelle valeur de V-test le marker est présent dans ce tableau ?

mettre la p-value , non ? }}
\label{tab:markers}
\end{table*}

@@

The aim here is to validate the hypothesis that the removal of i-markers and ni-markers have an impact on the perceived pleasantness. To do so, we perform a an Analysis of Variance. We consider $\mathcal{A}_{subject}$ as a dependent variable \jfp{je crois que la défintion de la variable dépendante, Asubject, n'est pas la bonne

A subject a a été définit avant comme "Asubject is
the average pleasantness for each participant"
l'ANOVA à deux facteurs à mesures répétées ne porte pas sur la valeur moyenne mais sur la valeur de l'évaluation pour une scene donnée

il faut donc définir une notation appropriée
}, and as independent variables, the type of environment (i/ni), and the presence/absence of markers (km/rm). As each subject evaluated the whole set of scenes, a two factors repeated measure ANOVA is used to check the validity of the hypothesis that there exist a significant difference of perceived pleasantness between km-scenes and rm-scenes. The two independent variables are considered as \emph{within-subject} factors. The factors being of only two levels each (type: i/ni,  marker: km/rm), the sphericity hypothesis does not need to be checked. \emph{Post hoc} analyses are done using the Tukey-Kramer procedure.

@@

The \emph{post hoc} analysis exhibits significative differences between all the groups of observations, notably between the i/km-scenes ($p<0.05$) and the i/rm-scenes ($p<0.01$). \jfp{je ne comprends pas les 2 p-value

il ne peut y a voir qu'une seule p-value lorsqu'on compare deux conditions expérimentales
}

>>>>

\documentclass[preprint,12pt]{elsarticle}
\usepackage{url}
\usepackage[utf8]{inputenc}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{caption}
\usepackage{sectsty}
\usepackage{amsmath}
\usepackage{color}

@@

\newcommand{\gl}[1]{\textcolor{magenta}{GL : #1}}
\newcommand{\nm}[1]{\textcolor{magenta}{#1}}
\newcommand{\ml}[1]{\textcolor{red}{#1}}
\newcommand{\jfp}[1]{\textcolor{blue}{JFP : #1}}
\newcommand{\ie}{\emph{i.\,e.}}
\newcommand{\Ie}{\emph{I.\,e.}}
\newcommand{\cf}{cf.}
\newcommand{\Cf}{Cf.}
\newcommand{\vs}{\emph{vs.}}
\newcommand{\Vs}{\emph{Vs.}}
\newcommand{\al}{\emph{et~al.}}
\newcommand{\eg}{\emph{e.\,g.}}
\newcommand{\Eg}{\emph{E.\,g.}}
\newcommand{\etc}{\emph{etc.}}
\newcommand{\myfloatalign}{\centering}

@@

\begin{itemize}
\item \emph{checking the pleasantness of the simulated sound scene}: to check that the emotional difference produced by the i- and ni-scenes is also present in terms of perceived pleasantness, we look at the existence of differences between the average $\mathcal{A}_{scene}$ and the average $\mathcal{A}_{subject}$ for each type of scenes (i or ni). The significance is evaluated by a Student test on two independent populations for $\mathcal{A}_{scene}$ and by a Student test on two paired populations for $\mathcal{A}_{subject}$;
\jfp{ce qu'est Ascene dans ce paragraphe n'est pas clair :
A scene est défini comme la moyenne des scores d'une scène donnée pour tous les participants.
le test, lui, porte sur la moyenne des Ascene pour les deux catégories
meme chose pour les sujets}
\item \emph{analysis of the structural features}: in order to evaluate whether the affective difference between  the i- and ni-scenes has a significative influence  on the structural features of the scenes \ie~ if significative differences exist between structural features and perceived pleasantness. The significance is measured with a Student test on two populations \jfp{je ne comprends pas bien ce test}
\item \emph{study of the influence of the structural features on the perceived pleasantness}: in order to evaluate a possible impact of the structural features on the perceived pleasantness, linear correlations between these features and $\mathcal{A}_{scene}$ are considered. The Pearson correlation coefficient is used for that purpose, on the basis of the standard methodology used in multidimensional studies;
\item \emph{analysis of the semantic features}: in order to determine if the affective difference between  the i- and ni-scenes has an influence on the scenes composition in terms of sound sources, or if specific classes are more frequently used to simulate a given environment (i or ni), a V-test is considered. Testing is done at every level of abstraction, and separately for texture and event classes. For each class $j$, and each environment type $k$ ($k=\lbrace i,ni\rbrace$), the value $V_{jk}$ is computed as follows:

@@

where $c$ is the number of available classes, $c_k$ the number of classes for a given type of environment $k$, $c_j$ the number of  $j$ classes, and $c_{jk}$ the number of $j$ classes used for an environment $k$. \jfp{il faudrait donner une référence pour ce test : bouquin Lebart, Morineau, Piron, STATISTIQUE EXPLORATOIRE MULTIDIMENSIONNELLE
peut être donner les bornes de variations de Vjk pour aider le lecteur à comprendre cet indicateur} \ml{y pas une ref en Anglais ?} The V-test tests the null hypothesis that the ratio $\frac{c_{jk}}{c}$ is not significatively different from the ratio $\frac{c_{jk}}{c_k}$. For a given environment $k$, and class $j$, if the hypothesis is rejected, the class $j$ is said to be typical with respect to environment $k$. The typical classes are called \textbf{sound markers};\jfp{on ne comprend pas bien l'hypothèse alternative du test. VOIR BOUQUIN LEBART page 184 ou bien article de Morineau sur le sujet Vjk est une estimation approchée de la valeur test}

@@

\begin{table*}[t]
 \setlength{\tabcolsep}{0.2pt}
 \centering
  {\renewcommand{\arraystretch}{0.9}
\begin{tabular}{c c c c}
Abstraction        & \multicolumn{2}{c}{Event sound markers} \\
level & i-scenes & ni-scenes \\
\hline
0  &                          & construction work (3.78)  \\
\hline
  & church bell  (4.5)             & car horn  (3.9) \\
1 & bicycle bell  (4.3)      & siren (3.9)\\
  & animal (4.2)              &       \\
   \hline
  & bird        (4.8)       & car horn  (4.0)\\
2 & church bell  (4.4)             & siren (4.0)\\
  & bicycle bell     (4.2)             &       \\
   \hline
  & bird song (4.8)        & klaxon  (4.1)\\
3 & church bell   (4.3)            & siren (4.0)\\
  & bicycle bell      (4.2)   &       \\
  & foot steps  (3.6)      &  \\
  &                           & \\
  & \multicolumn{2}{c}{Texture sound markers}      \\
  & i-scenes & ni-scenes \\
\hline
0 &     courtyard / park (4.1) &  construction (3.9)  \\
\hline
1 &     park (3.65)          &  crossroad (3.6)  \\
  &                          &  construction vehicule (3.3)  \\
\hline
2 &     park (3.64)          &  crossroad (3.56)  \\
\hline
\end{tabular}
}
\vspace{0.5mm}
\caption{Event and texture classes identified as sound markers. In each cell, markers are ranked in the decreasing order of V-test value.\jfp{a partir de quelle valeur de V-test le marker est présent dans ce tableau ? mettre la p-value , non ? }}
\label{tab:markers}
\end{table*}

@@

The aim here is to validate the hypothesis that the removal of i-markers and ni-markers have an impact on the perceived pleasantness. To do so, we perform a an Analysis of Variance. We consider $\mathcal{A}_{subject}$ as a dependent variable \jfp{je crois que la défintion de la variable dépendante, Asubject, n'est pas la bonne. A subject a a été définit avant comme "Asubject is
the average pleasantness for each participant". l'ANOVA à deux facteurs à mesures répétées ne porte pas sur la valeur moyenne mais sur la valeur de l'évaluation pour une scene donnée. il faut donc définir une notation appropriée}, and as independent variables, the type of environment (i/ni), and the presence/absence of markers (km/rm). As each subject evaluated the whole set of scenes, a two factors repeated measure ANOVA is used to check the validity of the hypothesis that there exist a significant difference of perceived pleasantness between km-scenes and rm-scenes. The two independent variables are considered as \emph{within-subject} factors. The factors being of only two levels each (type: i/ni,  marker: km/rm), the sphericity hypothesis does not need to be checked. \emph{Post hoc} analyses are done using the Tukey-Kramer procedure.

@@

The \emph{post hoc} analysis exhibits significative differences between all the groups of observations, notably between the i/km-scenes ($p<0.05$) and the i/rm-scenes ($p<0.01$). \jfp{je ne comprends pas les 2 p-value. il ne peut y a voir qu'une seule p-value lorsqu'on compare deux conditions expérimentales
}
